### 2024-02-06 README.Rmd
with data folder
remove inst/ext... and change function examples

### UK data

library(parlitools)

# prepare the data
# extract and join census and election data from parlitools package
census_11 <- parlitools::census_11 |> 
  select(-constituency_name,-constituency_type,-pano, -region, -country)
bes_2019 <- parlitools::bes_2019

elect_results <- left_join(bes_2019,census_11, by=c("ons_const_id"))

# download constituency boundaries for 2019 using the st_read() command below
# they come from: https://geoportal.statistics.gov.uk/datasets/ons::wpc-dec-2019-ultra-generalised-clipped-boundaries-uk
# which results in file: WPC_Dec_2019_UGCB_UK_2022_-5406762260244410508.geojson

# uk_map_download <- st_read(
#   "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/WPC_Dec_2019_UGCB_UK_2022/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson",
#   quiet = TRUE)

uk_map_download <- st_read(
  "inst/extdata/WPC_Dec_2019_UGCB_UK.shp",
  quiet = TRUE)

# only need the boundaries and the IDs for merging with parlitools data
uk<- uk_map_download |> 
  select(pcon19cd,geometry) |> 
  st_transform(crs=27700) |> 
  st_make_valid() # ensure valid line overlaps etc

# create dataframe for filtering-out speaker constituencies below
speakers <- data.frame(
  year = c(2017,2019),
  constituency_name = c("Buckingham","Chorley")
)

# join constituency polygons to parlitools data
# filter out Scotland and Northern Ireland
# filter out the speakers' constituencies
# calculate the desired explanatory variables
# make levels into factors for use with `mgcv`

df <- elect_results |> 
  left_join(uk, by=c("ons_const_id"="pcon19cd")) |> 
  # filter(!country %in% c("Scotland","Northern Ireland")) |> 
  filter(!country %in% c("Northern Ireland")) |> 
  # filter(!constituency_name %in% speakers$constituency_name) |>
  mutate(degree_educated = qual_level_4,
         health_not_good = health_fair + health_bad + health_very_bad,
         white = ethnicity_white,
         con_change = con_19 - con_17, # difference in %
         lab_change = lab_19 - lab_17,
         con_swing = (con_change - lab_change)/2, # Butler swing
         region = factor(ifelse(county == "Merseyside", "Merseyside", region)),
         county = factor(county)) |> 
  st_as_sf()

# give the empty speaker seats the mean value, just for demostration purposes here
df$con_swing[df$constituency_name %in% speakers$constituency_name] <- mean(df$con_swing,na.rm = TRUE)

# create region boundaries layer for mapping
regions <- df |> group_by(region) |> 
  summarise() |> 
  st_as_sf() |> 
  st_transform(crs=27700)

# create county boundaries layer for mapping
counties <- df |> group_by(county) |> 
  summarise() |> 
  st_as_sf() |> 
  st_transform(crs=27700)

# create outline layer for mapping
uk_outline <- df |> 
  summarise() |> 
  st_as_sf() |> 
  st_transform(crs=27700)

# filter out required independent variables and scale them
# then add back to the features which were not scaled
df_scaled <- df |> 
  st_drop_geometry() |> 
  select(degree_educated,
         health_not_good,
         white) |> 
  scale() |> 
  as.data.frame() |> 
  mutate(con_swing = df$con_swing,
         population = df$population,
         region = factor(df$region),
         county = factor(df$county),
         constituency_name = df$constituency_name)

# create a simple features (spatial) version of this with a geometry column
df_scaled_sf <- df_scaled |> 
  mutate(geometry = df$geometry) |> 
  st_as_sf() |> 
  st_transform(crs=27700)
```

@@

```{r, echo=FALSE}
df_scaled_sf$country <- case_when(df_scaled_sf$region == "Wales" ~ "Wales",
                                  df_scaled_sf$region == "Scotland" ~ "Scotland",
                                  TRUE ~ "England")
```

@@

```{r, fig.width=12, fig.height=10, echo=FALSE}
ggplot() + 
  geom_sf(data=df_scaled_sf, aes(fill=country), colour="white") + 
  # geom_sf_label(data=df_scaled_sf |> group_by(country) |> summarise(), aes(label=country),size=10) +
  geom_sf(data=df_scaled_sf |> filter(constituency_name == "Na h-Eileanan An Iar"), colour="#B20016", fill=NA, linewidth=0.7) + 
  geom_sf_label(data=df_scaled_sf |> filter(constituency_name == "Na h-Eileanan An Iar"),
                aes(label=constituency_name)) +
  geom_sf(data=df_scaled_sf |> filter(constituency_name == "Orkney and Shetland"), colour="#B20016", fill=NA, linewidth=0.7) + 
  geom_sf_label(data=df_scaled_sf |> filter(constituency_name == "Orkney and Shetland"),
                aes(label=constituency_name), nudge_x = -80000, nudge_y = -60000) +
    geom_sf(data=df_scaled_sf |> filter(constituency_name=="Ynys Mon"),colour="#B20016", fill=NA, linewidth=0.7) + 
    geom_sf_label(data=df_scaled_sf |> filter(constituency_name=="Ynys Mon"),
                  aes(label=constituency_name), nudge_x = -90000) +
    geom_sf(data=df_scaled_sf |> filter(constituency_name=="Isle Of Wight"),colour="#B20016", fill=NA, linewidth=0.7) + 
    geom_sf_label(data=df_scaled_sf |> filter(constituency_name=="Isle Of Wight"),
                  aes(label=constituency_name), nudge_x = 50000, nudge_y = -30000) +
  scale_fill_manual(values = c("antiquewhite4","#1C4769", "#24975E")) +
  # guides(fill = "none") + 
  theme_void()
```

@@

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = df_scaled_sf,
                   geom_col_name = "constituency_name",
                   remove_islands = T)
st_quickmap_nb(nbsf,
                    pointsize=0.05,
                    title = "st_bridges() contiguities",
                    subtitle = "no island constituencies\n(islands which remain are part of a contiguous constituency)")

@@

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = df_scaled_sf,
                   geom_col_name = "constituency_name",
                   link_islands_k = 2)
st_quickmap_nb(nbsf,
                    pointsize=0.05, 
                    title = "st_bridges() contiguities",
                    subtitle = "islands linked to k=2 nearest constituencies")

@@

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = df_scaled_sf,
                   geom_col_name = "constituency_name",
                   remove_islands = F,
                   link_islands_k = 2,
                   nb_structure = "matrix",
                   add_to_dataframe = T)
st_quickmap_nb(nbsf,
                    pointsize=0.05)

@@

```{r}
st_bridges(df = df_scaled_sf|> filter(region %in% c("Wales","South West")),
           geom_col_name = "constituency_name",
           link_islands_k = 2)  |> 
  st_manual_join_nb("Gower","St Ives")
```

@@

df_scaled_sf |> 
  filter(region == "Scotland") |> 
  mutate(nb = st_contiguity(df_scaled_sf$geometry[df_scaled_sf$region == "Scotland"]) |> 
           st_nb_lag_cumul(2)) |>  
  st_quickmap_nb()

@@

df_scaled_sf |> 
  filter(region == "Scotland") |> 
  st_bridges(geom_col_name = "constituency_name",
             link_islands_k = 2) |> 
  st_quickmap_nb()

@@

df_scaled_sf |> 
  filter(region == "Scotland") |> 
  st_bridges(geom_col_name = "constituency_name",
             link_islands_k = 2) |> 
  st_check_islands()
```

@@

df_scaled_sf |> 
  filter(region == "Scotland") |> 
  mutate(nb = st_contiguity(df_scaled_sf$geometry[df_scaled_sf$region == "Scotland"]) |> 
           st_nb_lag_cumul(2)) |> 
  st_manual_join_nb(47,9) |>
  st_manual_join_nb(47,51) |>
  st_manual_join_nb(47,55) |>
  st_manual_join_nb(51,9) |>
  st_manual_join_nb(51,47) |>
  st_quickmap_nb()

@@

prep_data <- st_bridges(df_scaled_sf, "constituency_name")

@@

prep_data2 <- st_bridges(df_scaled_sf, "constituency_name", nb_structure = "matrix")

@@

prep_data3 <- st_bridges(df_scaled_sf, "constituency_name") # decide upon the contiguities and add them to the df

>>>>

# ### UK data
# 
# library(parlitools)
# 
# # prepare the data
# # extract and join census and election data from parlitools package
# census_11 <- parlitools::census_11 |> 
#   select(-constituency_name,-constituency_type,-pano, -region, -country)
# bes_2019 <- parlitools::bes_2019
# 
# elect_results <- left_join(bes_2019,census_11, by=c("ons_const_id"))
# 
# # download constituency boundaries for 2019 using the st_read() command below
# # they come from: https://geoportal.statistics.gov.uk/datasets/ons::wpc-dec-2019-ultra-generalised-clipped-boundaries-uk
# # which results in file: WPC_Dec_2019_UGCB_UK_2022_-5406762260244410508.geojson
# 
# uk_map_download <- st_read(
#   "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/WPC_Dec_2019_UGCB_UK_2022/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson",
#   quiet = TRUE)
# 
# # only need the boundaries and the IDs for merging with parlitools data
# uk<- uk_map_download |> 
#   select(pcon19cd,geometry) |> 
#   st_transform(crs=27700) |> 
#   st_make_valid() # ensure valid line overlaps etc
# 
# # create dataframe for filtering-out speaker constituencies below
# speakers <- data.frame(
#   year = c(2017,2019),
#   constituency_name = c("Buckingham","Chorley")
# )
# 
# # join constituency polygons to parlitools data
# # filter out Scotland and Northern Ireland
# # filter out the speakers' constituencies
# # calculate the desired explanatory variables
# # make levels into factors for use with `mgcv`
# 
# df <- elect_results |> 
#   left_join(uk, by=c("ons_const_id"="pcon19cd")) |> 
#   # filter(!country %in% c("Scotland","Northern Ireland")) |> 
#   filter(!country %in% c("Northern Ireland")) |> 
#   # filter(!constituency_name %in% speakers$constituency_name) |>
#   mutate(degree_educated = qual_level_4,
#          health_not_good = health_fair + health_bad + health_very_bad,
#          white = ethnicity_white,
#          con_change = con_19 - con_17, # difference in %
#          lab_change = lab_19 - lab_17,
#          con_swing = (con_change - lab_change)/2, # Butler swing
#          region = factor(ifelse(county == "Merseyside", "Merseyside", region)),
#          county = factor(county)) |> 
#   st_as_sf()
# 
# # give the empty speaker seats the mean value, just for demostration purposes here
# df$con_swing[df$constituency_name %in% speakers$constituency_name] <- mean(df$con_swing,na.rm = TRUE)
# 
# # create region boundaries layer for mapping
# regions <- df |> group_by(region) |> 
#   summarise() |> 
#   st_as_sf() |> 
#   st_transform(crs=27700)
# 
# # create county boundaries layer for mapping
# counties <- df |> group_by(county) |> 
#   summarise() |> 
#   st_as_sf() |> 
#   st_transform(crs=27700)
# 
# # create outline layer for mapping
# uk_outline <- df |> 
#   summarise() |> 
#   st_as_sf() |> 
#   st_transform(crs=27700)
# 
# # filter out required independent variables and scale them
# # then add back to the features which were not scaled
# df_scaled <- df |> 
#   st_drop_geometry() |> 
#   select(degree_educated,
#          health_not_good,
#          white) |> 
#   scale() |> 
#   as.data.frame() |> 
#   mutate(con_swing = df$con_swing,
#          population = df$population,
#          region = factor(df$region),
#          county = factor(df$county),
#          constituency_name = df$constituency_name)
# 
# # create a simple features (spatial) version of this with a geometry column
# df_scaled_sf <- df_scaled |> 
#   mutate(geometry = df$geometry) |> 
#   st_as_sf() |> 
#   st_transform(crs=27700)

# df_scaled_sf <- uk_election
```

@@

```{r, echo=FALSE}
uk_election$country <- case_when(uk_election$region == "Wales" ~ "Wales",
                                  uk_election$region == "Scotland" ~ "Scotland",
                                  TRUE ~ "England")
```

@@

```{r, fig.width=12, fig.height=10, echo=FALSE}
ggplot() + 
  geom_sf(data=uk_election, aes(fill=country), colour="white") + 
  # geom_sf_label(data=uk_election |> group_by(country) |> summarise(), aes(label=country),size=10) +
  geom_sf(data=uk_election |> filter(constituency_name == "Na h-Eileanan An Iar"), colour="#B20016", fill=NA, linewidth=0.7) + 
  geom_sf_label(data=uk_election |> filter(constituency_name == "Na h-Eileanan An Iar"),
                aes(label=constituency_name)) +
  geom_sf(data=uk_election |> filter(constituency_name == "Orkney and Shetland"), colour="#B20016", fill=NA, linewidth=0.7) + 
  geom_sf_label(data=uk_election |> filter(constituency_name == "Orkney and Shetland"),
                aes(label=constituency_name), nudge_x = -80000, nudge_y = -60000) +
    geom_sf(data=uk_election |> filter(constituency_name=="Ynys Mon"),colour="#B20016", fill=NA, linewidth=0.7) + 
    geom_sf_label(data=uk_election |> filter(constituency_name=="Ynys Mon"),
                  aes(label=constituency_name), nudge_x = -90000) +
    geom_sf(data=uk_election |> filter(constituency_name=="Isle Of Wight"),colour="#B20016", fill=NA, linewidth=0.7) + 
    geom_sf_label(data=uk_election |> filter(constituency_name=="Isle Of Wight"),
                  aes(label=constituency_name), nudge_x = 50000, nudge_y = -30000) +
  scale_fill_manual(values = c("antiquewhite4","#1C4769", "#24975E")) +
  # guides(fill = "none") + 
  theme_void()
```

@@

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = uk_election,
                   geom_col_name = "constituency_name",
                   remove_islands = T)
st_quickmap_nb(nbsf,
                    pointsize=0.05,
                    title = "st_bridges() contiguities",
                    subtitle = "no island constituencies\n(islands which remain are part of a contiguous constituency)")

@@

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = uk_election,
                   geom_col_name = "constituency_name",
                   link_islands_k = 2)
st_quickmap_nb(nbsf,
                    pointsize=0.05, 
                    title = "st_bridges() contiguities",
                    subtitle = "islands linked to k=2 nearest constituencies")

@@

```{r, fig.width=12, fig.height=10}
nbsf <- st_bridges(df = uk_election,
                   geom_col_name = "constituency_name",
                   remove_islands = F,
                   link_islands_k = 2,
                   nb_structure = "matrix",
                   add_to_dataframe = T)
st_quickmap_nb(nbsf,
                    pointsize=0.05)

@@

```{r}
st_bridges(df = uk_election|> filter(region %in% c("Wales","South West")),
           geom_col_name = "constituency_name",
           link_islands_k = 2)  |> 
  st_manual_join_nb("Gower","St Ives")
```

@@

uk_election |> 
  filter(region == "Scotland") |> 
  mutate(nb = st_contiguity(uk_election$geometry[uk_election$region == "Scotland"]) |> 
           st_nb_lag_cumul(2)) |>  
  st_quickmap_nb()

@@

uk_election |> 
  filter(region == "Scotland") |> 
  st_bridges(geom_col_name = "constituency_name",
             link_islands_k = 2) |> 
  st_quickmap_nb()

@@

uk_election |> 
  filter(region == "Scotland") |> 
  st_bridges(geom_col_name = "constituency_name",
             link_islands_k = 2) |> 
  st_check_islands()
```

@@

uk_election |> 
  filter(region == "Scotland") |> 
  mutate(nb = st_contiguity(uk_election$geometry[uk_election$region == "Scotland"]) |> 
           st_nb_lag_cumul(2)) |> 
  st_manual_join_nb(47,9) |>
  st_manual_join_nb(47,51) |>
  st_manual_join_nb(47,55) |>
  st_manual_join_nb(51,9) |>
  st_manual_join_nb(51,47) |>
  st_quickmap_nb()

@@

prep_data <- st_bridges(uk_election, "constituency_name")

@@

prep_data2 <- st_bridges(uk_election, "constituency_name", nb_structure = "matrix")

@@

prep_data3 <- st_bridges(uk_election, "constituency_name") # decide upon the contiguities and add them to the df
